---
title: "PA1_template"
output: html_document
date: "Tuesday, October 14, 2014"
---
###Introduction

It is now possible to collect a large amount of data about personal movement using activity monitoring devices such as a Fitbit, Nike Fuelband, or Jawbone Up. These type of devices are part of the "quantified self" movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. But these data remain under-utilized both because the raw data are hard to obtain and there is a lack of statistical methods and software for processing and interpreting the data.

This assignment makes use of data from a personal activity monitoring device. This device collects data at 5 minute intervals through out the day. The data consists of two months of data from an anonymous individual collected during the months of October and November, 2012 and include the number of steps taken in 5 minute intervals each day.

The variables included in this dataset are:

    - steps: Number of steps taking in a 5-minute interval (missing values are coded as NA)
    
    - date: The date on which the measurement was taken in YYYY-MM-DD format
    
    - interval: Identifier for the 5-minute interval in which measurement was taken



```{r}
#Read data into R
data <- read.csv("c:/rWork/activity.csv", header=T,
                   stringsAsFactors = FALSE)

#Install ggplot2
library("ggplot2", lib.loc="~/R/win-library/3.1")

#Order data by date
dataord <- data[order(data$date, decreasing=FALSE), ]
#Set X and Y axis values
a <- c(1, 10750, 21500)
b <- c("1-Oct-2012", "31-Oct-2102", "30-Nov-2012")
c <- c(0, 3.5, 7)
d <- c("0", "450", "900")
#Sum steps by day
stepsXdate <- aggregate(x=dataord$steps, by=list(dataord$date), FUN="sum")
#Order file stepsXdate by date 
dataord2 <- stepsXdate[order(stepsXdate$x, decreasing=FALSE), ]
#Calculate the mean and median of the total number of steps taken per day
mn <- mean(stepsXdate$x, na.rm=T)
mn
md <- median(stepsXdate$x, na.rm=T)
md
#Create a histogram of steps taken per day
hist(stepsXdate$x, main = "Total Steps By Day", xlab="Days", ylab="Steps",
     xaxt="n", yaxt="n", breaks=61)
axis(1, at=a, labels=b)
axis(2, at=c, labels=d)
#Order data by interval
dataord3 <- data[order(data$interval, decreasing=FALSE), ]
 
plot(dataord3$interval, dataord3$steps, type="l", xlab= "Intervals",
     ylab= "Average Steps", col="blue" , lwd=2, main="Average Number of Steps")
#Find and print the interval with the maximum number of steps
maxSteps <- max(dataord3$steps, na.rm=T)
maxout <- subset(dataord3, steps==maxSteps, select=c(steps, date, interval))
maxout
#Calculate and print the total number of NA's
x <- is.na(dataord3$steps); y <- which(x); z <- length(y)
z






```



```{r, echo=TRUE}
aggmean <- tapply(dataord3$steps, dataord$interval, mean, na.rm=T)
taggmean <- mean(aggmean)
dataord3$steps[is.na(dataord3$steps)] <- taggmean
stepsXdateNA <- aggregate(x=dataord3$steps, by=list(dataord$date), FUN="sum")
mnNA <- mean(stepsXdateNA$x, na.rm=T)
mnNA
mdNA <- median(stepsXdateNA$x, na.rm=T)
mdNA
aa <- c(1, 23000, 46000)
bb <- c("1-Oct-2012", "31-Oct-2102", "30-Nov-2012")
cc <- c(0, 7.5, 15)
dd <- c("0", "23000", "46000")
hist(stepsXdateNA$x, main = "Total Steps By Day x NA's", xlab="Days", ylab="Steps",
     xaxt="n", yaxt="n", breaks=61)
axis(1, at=aa, labels=bb)
axis(2, at=cc, labels=dd)
dataord4 <- dataord3
dataord4$day <- "w"
dataord4$day <- ifelse(!weekdays(as.Date(dataord4$date)) %in% c("Saturday", "Sunday"),
       "weekday","weekend")
ggplot(dataord4,aes(interval,steps)) +geom_line() + facet_grid(.~day)
```


